Viscous Transport in Eroding Porous Media

### 2019-10-09 docs/transport/erosion_transport.tex
NM

Flow in porous media is further complicated when boundaries dynamically
evolve in response to the fluid flow. This coupling between the geometry
and the flow occurs, for example, in applications involving
melting~\citep{bec-vis1998, rycroft2016asymmetric, jambon2018singular,
favier2019rayleigh, morrow2019moving},
dissolution~\citep{kan-zha-che-he2002, mac2015shape, moo2017,
wykes2018self}, deposition~\citep{joh-eli1995, hewett2018modelling},
biofilm growth~\citep{tan-val-wer2015}, and crack
formation~\citep{cho2019crack}. We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical, hydrological, and
industrial applications~\citep{berhanu2012shape, hewett2017evolution,
lachaussee2018competitive, lopez2018cfd, allen2019sde, amin2019role}.
When a porous geometry erodes, certain qualitative characteristics are
unveiled that affect transport through the geometry.  For example, an
eroded geometry may contain channels of high porosity, which, though few
in number and modest in volume fraction, transmit a large portion of the
flux~\citep{qua-moo2018}. This arrangement results in velocities that
vary over several orders of magnitude~\citep{all-hea-lab-rei2002}.
Moreover, channelization creates heterogeneous and anisotropic medium
properties, which affect the transport of tracers such as
contaminants~\citep{cve-che-wen1996, dag1987, kon-bre1978} and
heat~\citep{nil-sto1990, ree-sto1995}.

This study consists of two main undertakings: first, high-fidelity
simulations of eroding porous media, and, second, characterization of
tracer transport through the resulting eroded geometries. Our modeling
efforts build on previous work~\citep{ris-moo-chi-she-zha2012,
moore2013self, moo2017}, in particular recent numerical methods
developed to simulate erosion in the Stokes-flow
regime~\citep{qua-moo2018}. We, however, make key improvements to the
numerical methods from~\citep{qua-moo2018} to enable simulations of more
realistic, dense suspensions of eroding grains
(figures~\ref{fig:Eroding50vel} and~\ref{fig:Eroding50vort}).  Then, to
characterize transport through these configurations, we examine
coarse-grained variables through statistical analysis of tracer
trajectories.

@@

Owing to the scales present in groundwater flow~\citep{bea1972}, we model
the hydrodynamics with the two-dimensional incompressible Stokes
equations.  Meanwhile, individual grains erode at a rate proportional to
the hydrodynamic shear stress~\citep{wan-fel2004,
ris-moo-chi-she-zha2012, moore2013self, par-izu2000}.  Since the
vorticity and the shear stress are equivalent on solid boundaries, we
also compute the fluid vorticity for visualization purposes
(figure~\ref{fig:Eroding50vort}).  A variety of numerical methods for
the fluid solver are available including lattice-gas or lattice
Boltzmann~\citep{kop-kat-tim1996, dar-mcc1998}, finite
volumes~\citep{suo-liu-gan2019, den-ica-hid2018,
sie-ili-pri-riv-gua2019}, finite differences~\citep{leb-ded-dav-bou2007,
knu-car2005}, and cellular automata~\citep{rot1988}.  However, simulating
erosion requires a method that is high-order, can resolve
nearly-touching arbitrarily shaped grains, and can accurately compute
derivatives of the velocity field.  Since the governing equations are
linear and homogeneous, we address these challenges by converting the
fluid equations to a boundary integral equation (BIE).

>>>>

Flow in porous media is further complicated when boundaries dynamically
evolve in response to the fluid flow. This coupling between the geometry
and the flow occurs, for example, in applications involving
melting~\citep{bec-vis1998, rycroft2016asymmetric, jambon2018singular,
favier2019rayleigh, morrow2019moving},
dissolution~\citep{kan-zha-che-he2002, mac2015shape, moo2017,
wykes2018self}, deposition~\citep{joh-eli1995, hewett2018modelling},
biofilm growth~\citep{tan-val-wer2015}, and crack
formation~\citep{cho2019crack}. We focus on erosion, a fluid-mechanical
process that is prevalent in many geophysical, hydrological, and
industrial applications~\citep{berhanu2012shape, hewett2017evolution,
lachaussee2018competitive, lopez2018cfd, allen2019sde, amin2019role}.
When a porous medium erodes, certain qualitative characteristics are
unveiled that affect transport through the geometry.  For example, an
eroded geometry may contain channels of high porosity, which, though few
in number and modest in volume fraction, transmit a large portion of the
flux~\citep{qua-moo2018}. This arrangement results in velocities that
vary over several orders of magnitude~\citep{all-hea-lab-rei2002}.
Moreover, channelization creates heterogeneous and anisotropic medium
properties, which affect the transport of tracers such as
contaminants~\citep{cve-che-wen1996, dag1987, kon-bre1978} and
heat~\citep{nil-sto1990, ree-sto1995}.

This study consists of two main undertakings: first, high-fidelity
simulations of eroding porous media, and, second, characterization of
tracer transport through the resulting eroded geometries. Our modeling
efforts build on previous work~\citep{ris-moo-chi-she-zha2012,
moore2013self, moo2017}, in particular recent numerical methods
developed to simulate erosion in the Stokes-flow
regime~\citep{qua-moo2018}. We, however, make key improvements to the
numerical methods \nick{(I killed the reference here because I think its redundant)} to enable simulations of more
realistic, dense suspensions of eroding grains
(figures~\ref{fig:Eroding50vel} and~\ref{fig:Eroding50vort}).  Then, to
characterize transport through these configurations, we examine
coarse-grained variables through statistical analysis of tracer
trajectories.

@@

Owing to the scales present in groundwater flow~\citep{bea1972}, we model
the hydrodynamics with the two-dimensional incompressible Stokes
equations.  Meanwhile, individual grains erode at a rate proportional to
the hydrodynamic shear stress~\citep{wan-fel2004,
ris-moo-chi-she-zha2012, moore2013self, par-izu2000}.
\nick{We compute the vorticity in the fluid bulk since, on solid boundaries, vorticity reduces to shear and thus provides a convenient way to simultaneously visualize local erosion rates and changes in the surrounding flow} (figure~\ref{fig:Eroding50vort}).  
A variety of numerical methods for
the fluid solver are available including lattice-gas or lattice
Boltzmann~\citep{kop-kat-tim1996, dar-mcc1998}, finite
volumes~\citep{suo-liu-gan2019, den-ica-hid2018,
sie-ili-pri-riv-gua2019}, finite differences~\citep{leb-ded-dav-bou2007,
knu-car2005}, and cellular automata~\citep{rot1988}.  However, simulating
erosion requires a method that is high-order, can resolve
nearly-touching arbitrarily shaped grains, and can accurately compute
derivatives of the velocity field.  Since the governing equations are
linear and homogeneous, we address these challenges by converting the
fluid equations to a boundary integral equation (BIE).
